Knowledge \& the web: exercise session 6

### 2015-11-22 exercise_6/ex6.tex
added tabular
Update ex6.tex

\section{Rule learning}
One model we tried was to learn classification rules from the training data. Rule learning algorithms come with different options for several aspects of the algorithms. In the exercise session we experimented with two of these.
\par First, for choosing the conditions of a classification rule, there are two options: one is to add to the rule the condition which increases the accuracy the most, another is to add to the rule the condition which increases the information gain the most. The first option resulted in a rule set consisting of 363 rules which clearly overfitted the training data.
The second option resulted in this rule set:
\begin{itemize}

\item if Sex = male and Fare $\leq$ 15.173 then 0  (268 / 35)
\item if Sex = female and Fare $\textgreater$ 48.200 then 1  (4 / 76)
\item if Sex = male and Cabin = ? then 0  (0 / 0)
\item if Sex = male then 0  (152 / 62)
\item else 1  (61 / 121)

\end{itemize}
which is correct for 617 out of 779 training examples.
When evaluated on the test set, we got an accuracy of ?\\
A second aspect of rule learning is the desired pureness, which is the minimum ratio of the major class in a covered subset in order to consider the subset pure. The standard value for this is 0.9. We experimented with this value set to 0.8 and 1.
For the desired pureness set to 0.8 and the information criterion we got the following rule set:
\begin{itemize}

\item if Sex = male then 0  (420 / 97)
\item else 1  (64 / 189)

\end{itemize}
which is correct for 609 out of 770 training examples.
When evaluated on the test set, we got an accuracy of ?.
For the desired pureness set to 1 and the information gain criterion we got the following rule set:
\begin{itemize}

\item if Sex = male and Fare $\leq$ 15.173 then 0  (268 / 35)
\item if Sex = female and Fare $\textgreater$ 48.200 then 1  (4 / 76)
\item if Sex = male and Cabin = ? then 0  (0 / 0)
\item if Sex = male then 0  (152 / 62)
\item else 1  (61 / 121)

\end{itemize}
which is the same as the first rule set, which was obtained with the standard value for the desired pureness.\\
Next to these two, there are three other parameters for rule learning: sample ratio (which specifies the sample ratio of training data used for growing and pruning), minimal prune benefit (which specifies the minimum amount of benefit which must be exceeded over unpruned benefit in order to be pruned) and use local random seed (which indicates if a local random seed should be used for randomization). We did not experiment with these parameters, but kept the standard values for these, which were 0.9, 0.25 and no respectively.\\
\\
From these results we can clearly see that Sex, Fare and SibSp are features which are important for this task. That Sex and Fare were important features was to be expected. When the Titanic collided with an iceberg, women and children and people of the higher ranks were evacuated first. This is reflected in the data as we see that women and people which payed a higher fare are more likely to have survived. That the number of siblings and spouses is also an important feature can be explained as follows. People in need want to stay close to their loved ones at all costs. But how larger a group, how more difficult it was to get on a lifeboat together. As such, how more siblings and spouses someone had aboard, how less likely that person was to survive.

>>>>

\section{Rule learning}
One thing we tried was to learn classification rules from the training data.
Rule learning algorithms come with different options for several aspects of the algorithms. In the exercise session we experimented with two of these.\\
First, for choosing the conditions of a classification rule, there are two options: one is to add to the rule the condition which increases the accuracy the most, another is to add to the rule the condition which increases the information gain the most.\\
The accuracy criterion option resulted in the following rule set:
\begin{itemize}

\item if Parch > 5.500 then 0  (1 / 0)
\item if Fare = >10 and Sex = male then 0  (170 / 25)
\item if Sex = female then 1  (63 / 197)
\item else 0  (171 / 61)

\end{itemize}
%which is correct for 539 out of 688 training examples.
The information gain criterion option resulted in this rule set:
\begin{itemize}

\item if Sex = male and Fare = >10 then 0  (170 / 25)
\item if Sex = female then 1  (64 / 197)
\item if Fare = 10-20 and Parch ≤ 0.500 then 0  (66 / 7)
\item else 0  (111 / 54)

\end{itemize}
%which is correct for 544 out of 694 training examples.
A second aspect of rule learning is the desired pureness, which is the minimum ratio of the major class in a covered subset in order to consider the subset pure. The standard value for this is 0.9. We experimented with this value set to 0.8 and 1.
For the desired pureness set to 0.8 and the accuracy criterion we got the following rule set:
\begin{itemize}

\item if Parch > 5.500 then 0  (1 / 0)
\item if Fare = >10 then 0  (189 / 47)
\item if Sex = female then 1  (44 / 175)
\item else 0  (171 / 61)

\end{itemize}
%which is correct for 536 out of 688 training examples.
This rule set is very similar to the rule set obtained with the accuracy criterion option and desired pureness of 0.9. The only difference is that in the second rule the \textit{Sex = male} condition now disappeared from the condition part. 
For the desired pureness set to 0.8 and the information gain criterion we got the following rule set:
\begin{itemize}

\item if Sex = male then 0  (360 / 93)
\item else 1  (57 / 177)

\end{itemize}
%which is correct for 537 out of 687 training examples.
When comparing this rule set with the rule set obtained for the information gain criterion with the desired pureness set to 0.9, we see that this rule set is a great simplification of the other.
For the desired pureness set to 1 and the accuracy criterion we got exactly the same rule set as when setting the desired pureness to 0.9 for that criterion. The same happened for the desired pureness set to 1 and the information gain criterion, which also resulted in the same rule set as when setting the pureness to 0.9 and keeping the information criterion.\\
Next to the criterion and desired pureness, there are three other parameters for rule learning: sample ratio (which specifies the sample ratio of training data used for growing and pruning), minimal prune benefit (which specifies the minimum amount of benefit which must be exceeded over unpruned benefit in order to be pruned) and use local random seed (which indicates if a local random seed should be used for randomization). We did not experiment with these parameters, but kept the standard values for these, which were 0.9, 0.25 and no respectively.\\
\\
From these results we can clearly see that Sex, Fare and Parch are features which are important for this task. That Sex and Fare were important features was to be expected. When the Titanic collided with an iceberg, women and children and people of the higher ranks were evacuated first. This is reflected in the data as we see that women and people which payed a higher fare are more likely to have survived. That the number of parents and children is also an important feature can be explained as follows. People in need want to stay close to their loved ones at all costs. But how larger a group, how more difficult it was to get on a lifeboat together. Also, little children and old people need much more assistence to even reach a lifeboat. As such, how more children and parents someone had aboard, how less likely that person was to survive.
